---
title: "Inclass_ch6_anova"
output: html_document
date: "2024-02-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Install packages if not already installed
packages <- c("nlme", "lme4", "multcomp", "emmeans", "agricolae", "multcompView", "ggplot2", "lattice", "ggpubr", "plotly", "car", "agricolae","emmeans")
new_packages <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)

# Load the libraries
#library(nlme)     # For linear and nonlinear mixed-effects models
#library(lme4)     # For linear mixed models and generalized linear mixed models
# No need to load a library for aov() and lm() as they are part of base R
library(car)
library(emmeans)
library(multcomp)
library(multcompView) # For conducting post-hoc tests and general hypotheses testing
library(emmeans)  # For post-hoc analysis and estimation of marginal means
library(agricolae) # For multiple comparisons and group separations

library(ggplot2)  # For data visualization
library(lattice)  # For multi-panel graphics
library(ggpubr)   # For creating publication-ready plots
library(plotly)   # For interactive visualizations

# Note: This script assumes a connection to the internet to install packages.
# After installation, packages are loaded into the R session for use.

```

#table 4.4: Problem 4 pg 101
The experiment was run to compare the effects of auditory and visual cues on 
speed of response of a human subject. A personal computer was used to present 
a “stimulus” to a subject, and the reaction time required for the subject to 
press a key was monitored. The subject was warned that the stimulus was 
forthcoming by means of an auditory or a visual cue. The experimenters were 
interested in the effects on the subjects’ reaction time of the auditory and 
visual cues and also in different elapsed times between cue and stimulus. Thus,
there were two different treatment factors: “cue stimulus” at two levels 
“auditory” or “visual,” and “elapsed time between cue and stimulus” at three 
levels “five,” “ten,” or “fifteen” seconds. This gave a total of six treatment
combinations, which can be coded as:

```{r}
# Recreating the data frame from the provided image
rt_df <- data.frame(
  rep = c(1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2, 3),
  trt = factor(rep(c("aud", "aud", "aud", "vis", "vis", "vis"), each = 3)),
  time = rep(c(5, 10, 15), times = 6),
  respon_time_sec = c(0.204, 0.17, 0.181, 0.167, 0.182, 0.187,
                      0.202, 0.198, 0.236, 0.257, 0.279, 0.269,
                      0.283, 0.235, 0.26, 0.256, 0.281, 0.258)
)
rt_df$time <- factor(rt_df$time)
print(rt_df)
## Summarize the data to QC
summary(rt_df)

## Check the structure of the data frame
str(rt_df)

# Boxplot to explore response times across treatment groups
boxplot(respon_time_sec ~ trt, data = rt_df, main = "Response Time by Treatment",
        xlab = "Treatment", ylab = "Response Time (sec)", col = "lightblue")

# Boxplot to explore response times across time groups
boxplot(respon_time_sec ~ time, data = rt_df, main = "Response Time by Time",
        xlab = "Time", ylab = "Response Time (sec)", col = "lightblue")
```

```{r}
# Two-way ANOVA
aov_res <- aov(respon_time_sec ~ trt * time, data = rt_df)

# Calculate standardized residuals
aov_res_std_res <- rstandard(aov_res)

# Calculate fitted (predicted) values
aov_res_fitted <- fitted(aov_res)

# Base R plot for standardized residuals vs. fitted values
plot(aov_res_fitted, aov_res_std_res, xlab = "Fitted Values", ylab = "Standardized Residuals",
     main = "Standardized Residuals vs. Fitted Values")
abline(h = 0, col = "red")

# Create a data frame for plotting
residuals_data <- data.frame(StandardizedResiduals = aov_res_std_res, FittedValues = aov_res_fitted)

# ggplot2 for standardized residuals vs. fitted values
ggplot(residuals_data, aes(x = FittedValues, y = StandardizedResiduals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red") +
  labs(x = "Fitted Values", y = "Standardized Residuals",
       title = "Standardized Residuals vs. Fitted Values") +
  theme_minimal()

# Normality check with Shapiro-Wilk test
shapiro.test(rt_df$respon_time_sec)

# Homogeneity of variances with Levene's Test
leveneTest(respon_time_sec ~ trt * time, data = rt_df)
```


```{r}
summary(aov_res)
# Post-hoc tests
# Fisher's LSD
fisher_lsd <- LSD.test(res.aov, "trt", p.adj = "none")
print(fisher_lsd$groups)

# Tukey's HSD
tukey_hsd <- TukeyHSD(res.aov, "trt:time", conf.level = 0.95)
print(tukey_hsd)
```

```{r}
# Data visualization
# Interaction plot
interaction.plot(x.factor = rt_df$time, trace.factor = rt_df$trt, 
                 response = rt_df$respon_time_sec,
                 main = "Interaction Plot", xlab = "Time", ylab = "Response Time (sec)",
                 trace.label = "Treatment", fixed = TRUE)

# Boxplot of response times by treatment and time
ggplot(rt_df, aes(x = interaction(trt, time), y = respon_time_sec, fill = trt)) +
  geom_boxplot() +
  labs(title = "Response Times by Treatment and Time", x = "Treatment-Time Combination", y = "Response Time (sec)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Save the plot if needed
ggsave("boxplot_treatment_time.png", width = 10, height = 6, dpi = 300)
```

#In Class Work Data 1 - We do this one
```{r}
# Create the data frame for a 2 x 3 factorial design with 4 replicates
set.seed(42)  # Setting seed for reproducibility

# Define levels for Factor A and Factor B
factor_a_levels <- c("A1", "A2")
factor_b_levels <- c("B1", "B2", "B3")

# Create a data frame using expand.grid to get all combinations of the factorial design
factorial_design <- expand.grid(
  FactorA = factor_a_levels,
  FactorB = factor_b_levels,
  Rep = 1:4
)

# Generate response variable with interaction
# We'll simulate some data that shows an interaction between the two factors
factorial_design$Response <- with(factorial_design, 
                                  rnorm(nrow(factorial_design), 
                                        mean = ifelse(FactorA == "A1", 10, 15) +
                                          ifelse(FactorB == "B1", 0, ifelse(FactorB == "B2", 5, 10)), 
                                        sd = 1))

```

```{r}
summary(factorial_design)
factorial_plot <- aov(Response ~ FactorA * FactorB, data = factorial_design)
plot(factorial_plot)

model <- lm(Response ~ FactorA * FactorB, data = factorial_design)

# Normality test
shapiro.test(rstandard(model))
# I think this passes normality test?

```


#Inclass Data 2
```{r}
# Set the seed for reproducibility
set.seed(42)

# Define the levels for Factor A and Factor B
factor_a_levels1 <- c("A1", "A2")
factor_b_levels1 <- c("B1", "B2", "B3")

# Create a data frame with all combinations of Factor A and Factor B for 4 replicates
factorial_design1 <- expand.grid(
  FactorA1 = factor_a_levels1,
  FactorB1 = factor_b_levels1,
  Rep = 1:4
)

# Generate response variable with a strong interaction effect
# We will make the response for the combination of A2 and B3 much higher to emphasize interaction
factorial_design1$Response <- with(factorial_design1, 
                                  5 + 
                                  (as.numeric(FactorA1)-1) * 5 +
                                  (as.numeric(FactorB1-1) * 2 +
                                  (as.numeric(FactorA1)-1) * (as.numeric(FactorB1)-1) * 5 +
                                  rnorm(nrow(factorial_design1), mean = 0, sd = 1))


```

```{r}
summary(factorial_design1)
factorial_plot1 <- aov(Response1 ~ FactorA1 * FactorB1, data = factorial_design)
plot(factorial_plot1)
```


```



